A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-10-03 paper/paper.tex
paper: MAJ

@@

\subsection{Fibered versus indexed representation}
The fibered representation can be illustrated as:

@@

The indexed representation can be illustrated as:

@@

One way to obtain one construction from the other is via the \emph{Grothendieck construction}, although this is not relevant to our work.

@@

\section{Reasoning about sets in type theory}
Since proof irrelevance is an inherent part of set theory and first-order logic, and proof irrelevance is \emph{extensional} in type theory, we work in a universe known as \texttt{HSet}, which we describe shortly.

\subsection{Unicity of identity proofs}
UIP is a flavor of proof-irrelevance:

\subsection{HSet}
\texttt{HSet} provides us a restricted setting of UIP, and our implementation of \texttt{HSet} is a straightforward packaging of two fields:

@@

\renewcommand*{\lab}{tab:faces}
\input{tab-faces.tex}

@@

\renewcommand*{\lab}{tab:coh}
\input{tab-coh.tex}

>>>>

\subsection{Generalization to \texorpdfstring{$\nu$}{ν}-sets}
Let us call $\nu$-sets, the straightforward generalization of augmented semi-simplicial sets and semi-cubical sets to an arbirary arity, where morphisms are sequences of elements of an arbirary set $\nu$, extended with $\kstar$. Then, we observe:

@@

\subsection{Homotopy type theory}
Homotopy type theory (HoTT) is a logical foundation based on Martin-Löf type theory and the univalence axiom~\cite{hottbook}. HoTT is based on the notion of a basic \textsf{Type}, and the notion of sets can be recovered as \textsf{HSet}, where \texttt{HSet} is a straightforward packaging of two fields:

\textsf{HSet} provides us a form of unicity of identity of proofs (UIP), which is a form of proof irrelevance:

\subsection{Fibered versus indexed representation}
It is well-known in set theory, that families of small sets can be represented in a fibered or indexed way. In the fibered way, a family over $T$ is represented a pair of a set $S$, and a function $S \rightarrow T$. The function classifies the original set into fibers. In contrast, in the indexed way, a family over $T$ is represented as a functional relation with domain $T$, which itself, by replacement axiom, is a set.

In category theory, the correspondence between fibered and indexed way, manifests as the correspondence between Grothendieck fibrations and pseudofunctors, of which a particular instance is the correspondence between discrete fibrations and presheaves, or even between discrete fibrations over sets and families of sets.

In HoTT, the latter correspondence can be expressed, for $T: \textsf{Type}$, as an equivalence between $\Sigma S: \textsf{Type}\;.\;S \rightarrow T$ and $T \rightarrow \textsf{Type}$, which can be proved using univalence. This correspondence also holds on \textsf{HSet}, for $T: \textsf{HSet}$:

\begin{equation*}
  \Sigma S: \textsf{HSet}\;.\;S\rightarrow T \simeq T \rightarrow \textsf{HSet}
\end{equation*}

Let us now come back to cubical sets. A cubical set can be represented as a sequence of fibrations, together with appropriate coherence conditions:

@@

We can iteratively apply the above correspondence to obtain an indexed representation of cubical sets:

@@

The focus of our work has been to formally define such an indexed presentation of not just semi-cubical, but more generally, $\nu$-sets.

@@

@@

\renewcommand*{\lab}{tab:faces}
\input{tab-faces-ett.tex}

@@

\renewcommand*{\lab}{tab:coh}
\input{tab-coh-ett.tex}
